Debugging by \textit{lastChange}

### 2010-10-27 trunk/papers/pldi2011/sigplanconf-template.tex

If developer asks for the last change of variable \texttt{foo} at the 
reproduction point in a program, debugger calls 
\texttt{setVariableChangeHook()} with the variable's defining
block and name as parameters and re-executes the
program. Whenever \texttt{foo} changes and the callback function is to
be called, debugger first calls \texttt{scopeId()} on the
variable's scope. Then it stores this scope id, the stack
frame locations, and other state values in scope at the call point.
Whenever the execution reaches the reproduction point the debugger
looks at the history of \texttt{foo} changes and finds the last
\texttt{foo} change with the same scope id as the variable's scope id
at the reproduction point (Figure ~\ref{fig:foo-changes2}).

@@

@@

\subsection{Execution reproduction}
Although execution reproduction is basicly should be provided by the
developer, we tried to devise some automatic way which reproduces the
execution. In \textit{Querypointer}, develooper can directly provide a
test case which reproduces the execution. It is also possible to use
two automatic reproduction option: record-replay or local-reproduction.

Two elements should be carefully considered in execution
reproduction. First, the initial state should be the same as previous
execution. Second, the similar actions and events should be applied to
the program during the execution. In Record-replay approach, debugger
stores the initial page url and user actions and during replay phase,
opens the same url and simulates user actions. In the local
reproduction approach, debugger calls the funtion on the bottom of
stack with the same parameters. The idea behind local-reproduction is
that in many web pages it is enough to replay the event to reproduce
the bug. While record-replay approach gives better warranty about the
same initial states, the local-reproudction gives us shorter
cycles. It is also possible to uses thes approaches in
combination. For example, debugger locally reproduce the execution and
if it couldn't find the answer, reproduce the whole execution by
record-replay.
% should we talk about counting breakpoint hits for stopping at the right place?

\subsection{Data collection}
So far, we talked about the data collected at change events for
\textit{lastChange} algorithm. However, if developer needs to explore
the program state at a given point, more data is required. It is
possible to postpone data collection until user requests. It is also
possible to collect the list of available
variable. \textit{Querypointer} collects data in one level depth. More
data will be collected upon the developer's request by re-execution.

We needed to build an interface to different execution points which
are presented at the same time. The current \textit{Querypointer}
interface resembles the script panel interface while it keeps the
history of navigation.

@@

>>>>

If developer asks for the last change of variable \texttt{foo} at the 
reproduction point in a program, debugger calls 
\texttt{setVariableChangeH-} \texttt{ook()} with the variable's defining
block and name as parameters and re-executes the
program. Whenever \texttt{foo} changes and the callback function is to
be called, debugger first calls \texttt{scopeId()} on the
variable's scope. Then it stores this scope id, the stack
frame locations, and other state values in scope at the call point.
Whenever the execution reaches the reproduction point the debugger
looks at the history of \texttt{foo} changes and finds the last
\texttt{foo} change with the same scope id as the variable's scope id
at the reproduction point (Figure ~\ref{fig:foo-changes2}).

@@

Here, we implicitly assumed that the object will be created at 
the same location in the next execution. So, what does happens if it
is not true? Well, once the execution reaches the reproduction point, it
reveals that the object has been created in different location. This time,
debugger re-executes the program considering both locations as possible 
object creation locations.

@@

\subsection{Re-Execution, reproduction point and data collection}
\textit{Querypointer} needs a test case to reproduce the
execution and conditions to correctly recognize the reproduction point. 
Although both elements can be directly provided by developer, \textit{Querypointer}
is also able to automatically create them from the first execution.

\textit{Querypointer} keeps track of breakpoint hits and user steppings. For example, 
if developer query \textit{lastChange} at the third hit of breakpoint \textit{b}, in
re-execution, third hit of breakpoint b recognized as the reproduction point.

To build a test case from one execution, two elements should be carefully 
considered. First, the initial state should be the same as the initial state
of execution. Second, similar actions and events should be applied to
the program during the execution. \textit{Querypointer} supports two mechanism
for automatic re-execution: local-reproduction, record-replay. The first one,
recalls the function at the bottom of call stack with the same parameters. The idea
behind this mechanism is that many bugs in web pages can be reproduced by firing
an event like clicking on a button. The second one, in the first execution, record phase, 
stores the initial page url and captures user actions. In the replay phase, it opens
the same url and applies user actions. The local reproduction mechanism provides
shorter re-execution cycles while the record-replay is more accurate about the
initial state.

In addition to the data collected at every change event for identifying the \textit{lastChange}
result, \textit{Querypointer} partially stores values in program state. If developer asks for 
some values which have not been stored, \textit{Querypointer} re-executes and collects the requested data.
There is a trade-off between the amount of data collected at every change event and the number of re-executions. 
By default, \textit{Querypointer} stores values up to one level depth in objects.

@@

\bibitem[Bond(2010)]{Bond2}
M.D. Bond, G.Z. Baker, S.Z. Guyer, and Z. Samuel. \newblock Breadcrumbs: efficient context sensitivity for dynamic bug detection analyses.
\newblock In \emph{Conference on Programming Language Design and Implementation(PLDI)},
June, 2010.
